---
title: "Make Lookup Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make Lookup Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(conflicted)
suppressMessages(conflict_prefer("filter", "dplyr"))

library(DOPE)
library(tibble)
library(dplyr)
library(stringr)
library(usethis)
```

The lookup table has information on drug category, class and synonyms.  It is made from the DEA and "No Slang" tables. The core of the table is the dea_factsheets table.  It is tweaked to simplify complex names

## Fix dea_brands

The dea brands file has:

+ mixed case
+ registered trademark symbols
+ The drugs *kadian* is together with *ms contin* as a single record

```{r}
dea_brands <- DOPE::dea_brands %>%
  rename("brand" = brands) %>%
  mutate(across(where(is.character), tolower)) %>%
  mutate(brand = str_remove_all(brand, "®")) %>%
  mutate(brand = case_when(brand == "kadianms-contin" ~ "kadian", # two names
                           TRUE ~ brand)) %>%
  bind_rows(c(category = "morphine", brand = "ms contin"))
```

## Fix DEA category and class

The fact sheet file has:

+ mixed case
+ ghb listed with a complex name
    + gamma-hydroxybutyric acid should appear as ghb in all "class" variables
+ Mdma listed with 3 names
    + ecstacy, molly and mdma should be appear as mdma in all "class" variables
+ peyote and mescalin grouped together
    + peyote and mescalin should appear as mescalin in all "class" variables
+ flakka is listed with a complex name
    + flakka should appear as flakka in all "class" variables

These two do not fit well

+	spice/ k2, synthetic marijuana
+	acetaminophen and oxycodone


```{r}
dea_class_cat <- DOPE::dea_factsheets %>%
  mutate(across(where(is.character), tolower)) %>%
  mutate(category =
           case_when(category == "peyote and mescaline" ~ "mescaline",
                     category == "ghb - gamma-hydroxybutyric acid" ~ "ghb",
                     category == "ecstasy or mdma (also known as molly)" ~ "mdma",
                     category == "flakka (alpha-pvp)" ~ "flakka",

                     TRUE ~ category))

```

## The DEA street name file has:

+ mixed case
+ fentanyl has a complex name
    + fentanyl should appear as fentanyl in all "class" variables


```{r}
dea_street_names <- DOPE::dea_street_names %>%
  mutate(across(where(is.character), tolower)) %>%
  mutate(category =
           case_when(category == "fentanyl and fentanyl derivatives"	~ "fentanyl",
                     TRUE ~ category))

```


```{r}
# the these drugs do not have fact sheets but they appear on the DEA

# DEA website (in slang)
missing <- tribble(
  ~category, ~class, ~fact_path,
  "acetaminophen and oxycodone", "narcotics (opioids)", "DEA slang",
  "alprazolam", "benzodiazepine", "DEA slang" ,
  "clonazepam", "benzodiazepine", "DEA slang"  ,
  "crack cocaine", "stimulants", "DEA slang",
  "hydrocodone", "narcotics (opioids)", "DEA slang",
  "marijuana concentrates", "cannabis", "DEA slang",
  "mushrooms", "hallucinogen", "DEA slang",
  "PCP", "hallucinogen", "DEA slang",
  "promethazine with codeine", "narcotics (opioids)", "DEA slang",
  "ritalin", "stimulants", "DEA slang",
  "synthetic cannabinoids", "designer drugs", "DEA slang",
  "synthetic cathinones", "stimulants", "DEA slang")

# extras from NoSlang
extraNewSlang <- tribble(
  ~category,  ~class, ~fact_path,
  "2cb", "hallucinogen", "noSlang Term",
  "alpha-ethyltryptamine", "hallucinogen", "noSlang Term",
  "alpha-methyltryptamine", "hallucinogen", "noSlang Term",
  "amobarbital", "depressant", "noSlang Term",
  "amyl nitrite", "inhalant", "noSlang Term",
  "dextromethorphan", "depressant", "noSlang Term",
  "diazepam", "benzodiazepine", "noSlang Term",
  "dimethyltryptamine", "hallucinogen", "noSlang Term",
  "gbl", "depressant", "noSlang Term",
  "isobutyl nitrite", "inhalant", "noSlang Term",
  "methcathinone", "stimulant", "noSlang Term",
  "methaqualone", "depressant", "noSlang Term",
  "nitrous oxide", "inhalant", "noSlang Term"
)
```




```{r}
categories <- data.frame(allCat = c(dea_brands$category,
                                   dea_street_names$category,
                                   dea_class_cat$category,
                                   missing$category,
                                   extraNewSlang$category)) %>%
  distinct() %>%
  arrange()
```

```{r}
library(sqldf)
lookup_df <- sqldf("select cc.class, a.category, a.syn synonym from
                (select b.category, b.brand as syn from dea_brands as b
                 union
                 select s1.category, s1.slang as syn from dea_street_names as s1
                 union
                 select s2.category, s2.brand as syn from dea_street_names as s2 where s2.brand <> NULL
                 union
                 select ns.drug, ns.street_name as syn from noslang_street_names as ns) as a
              left join dea_class_cat as cc on a.category = cc.category")

#use_data(lookup_df, overwrite = TRUE)
```

Several records do not have a class. 

```{r fix_missing_class, echo=FALSE}
na_lookup <- filter(lookup_df, is.na(class))
length(unique(na_lookup$category)) # 30 unique categories with no class
# *am = depressants
# *mine = hallucinogen as per DAN 3/29/21
# *2cb = hallucinogen
# *amt = hallucinogen 
# *trite = depressants
# *deine = narcotics (opioids)
# crack = stimulants

lookup_df2 <- lookup_df %>% 
  mutate(class = case_when(str_detect(category, "(?=.*am$)|(?=.*trite$)|(gbl)|(?=.*qualone$)||(?=.*ital$)") == TRUE ~ "depressants",
                           str_detect(category, "crack|(?=.*cathin)|(?=.*date$)|(ritalin)") == TRUE ~ "stimulants",
                           str_detect(category, "(?=.*amt$)|(?=.*2cb$)|(?=.*mine$)|(dmt)|(?=.*room)|(pcp)|(peyote)") == TRUE ~ "hallucinogen",
                           str_detect(category, "(?=.*deine$)|(?=.*one$)") == TRUE ~ "narcotics (opioids)",
                           str_detect(category, "(?=.*marijuana)|(?=.*cannab)") == TRUE ~ "cannabis",
                           category == "dextromethorphan" ~ "antitussives",
                           category == "nitrous oxide" ~ "inhalant",
                           TRUE ~ class
         )
)

#check missing again
na_lookup <- filter(lookup_df2, is.na(class))
lookup_df <- lookup_df2

```

Add data from IQVIA

```{r}
iqvia <- DOPE::iqvia %>%
  mutate(across(where(is.character), tolower)) 

lookup_df <- DOPE::lookup_df %>% 
  bind_rows(iqvia)

use_data(lookup_df, overwrite = TRUE)
```

