---
title: "Introduction to parse() and the lookup_*() functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{parse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(DOPE)
library(tidyverse)
library(stringr)
```

This package aims to parse out identifiable drug names given a corpus of text. By corpus of text, we assume that the data has already been imported into R.

## Data: drug_df
---
Throughout this vignette, we will employ a sample dataset - `drug_df` - that is intended to represent data collected from a clinical trial. The dataset contains 3 variables and 500 observations. 

```{r data, echo=TRUE}
dim(drug_df)

str(drug_df)

class(drug_df)
```
Note: `drug_df` is a simulated dataset and does not reflect any true clinical observations. 

## `parse()`
---

The `parse()` function is intended to extract identifiable drug names from a corpus of text such as, clinical trial data, social media, survey or interview transcription. `parse()` takes in one argument, the vector that contains the strings to be parsed.

Here is an example of some problematic records in the `drug_df` dataset that warrants the use of `parse()`

```{r messy}

drug_df %>% 
  filter(str_detect(textdrug, ",|;|and|\\/|=|\\(")) %>% 
  distinct(textdrug)

```

As you can see there are so many extraneous/problematic characters, multiple drugs in one record and several variations of the same drug (i.e. "bup/nx"). We assume that the user is solely interested in the drugs themselves, not information such as dosage and units. 

This messy data is exactly what `parse()` was designed for. 

```{r}
drug_names <- parse(drug_df$textdrug)

head(drug_names)
```

The resulting vector can then be passed on to the `lookup_*` functions to identify whether the input drug is a class or category or a synonym for other drugs in the same category. 

## `lookup_*`
---

### `lookup()`

This function relies on a comprehensive lookup table `lookup_df`. This dataframe contains 3 variables:  
- `class` =  Highest level classification e.g. "stimulant" or "narcotic (opiod)"
- `category` =  More specific level of classification e.g. "heroin", "opium" or "marijuana"
- `synonym` =  Common name or street slang for specific drug name e.g. "china", "dope" or "weed"

```{r lookup_df, echo=TRUE}
dim(lookup_df)

str(lookup_df)

class(lookup_df)
```

The purpose of this function is to return any possible matches to the lookup_df, which is a comprehensive dataframe consisting of all drug classes, categories and synonyms. It serves as a source or helper function to many of the other more specific function in the package. The idea is to match any possible columns with a the single word, a list of separate words or a vector passed as an argument. The dataframe returned will consist of the lookup_df match as well as the `original_word` that was the source of the match. 

Here is an example of a common search done using `lookup`. 

```{r}
results <- lookup("crack", "pizza", "marijuana")
head(results)
```

You can see that the dataframe returned could be vast in its matches, and that the other more specific functions, below, might be of more use depending on one's needs. 

---

### `compress_lookup()`

This function takes in one argument: the table returned from a search using the `lookup` function. The purpose of this function is to narrow down the results to a more specific dataframe consisting of only relevant values, such as `class` and/or `category` depending on the user's selection. `compress_lookup` returns, by default, `original_word`, `class` and `category`. 


If a researcher wanted to determine the main classes of drugs being used by the patients of a clinical study, they might pass a large vector of substances from clinical notes taken in a study to the `lookup` function, then filter them down to only return the datafram of classes relevant to their needs.

Here is an example of a common search done using `compress_lookup`. 

```{r}
results <- lookup("crack", "pizza", "marijuana")
filtered_df <- compress_lookup(results, compressOriginalWord = TRUE, compressCategory = TRUE)
head(filtered_df)
```

The resulting dataframe is a short list of only the relevant information needed. 

---

### `lookup_syn()`

The purpose of this function is to find all possible synonyms of, primarily, a slang/street name of a commonly abused drug. Though searching for a drug class or category with `lookup()` will also return common synonms, this function makes searching specifically for synonyms explicit by taking just one argument: `drug_name`. The function will then determine the category of the slang term (drug_name) and return all synonyms that share that category. 

Here is an example of a common search done using `lookup_syn`. 

```{r}
results <- lookup_syn("china")
head(results)
```

The resulting dataframe contains a moderate list of terms that are synonyms of the `drug_name` given as determined by sources such as the DEA, FDA and other publicly available resources. 
